---
title: "Simulations"
output: html_notebook
---

#Does IOS accounts for pleiotropic effect indeed?
This simulation is designed to see whether decreased level of heterogeneity in MR analysis is because IOS score accounts for pleiotropic effect in MR model or just by chance.


```{r}
library(mr.ios)
library(parallel)
library(tidyverse)
```

Here we used an example of BMI and CHD for demonstration. IOS score was generated for each SNP (n=79) and used as an extra weight in IVW Radial MR model. Two different types of IOS were used: 

* IOS1: A summary of the SNP r2 with the other traits (r2_gu)
* IOS2: A summary of the ratio of r2_gu / r2_gx, where r2_gx is the variance explained by the SNP on the exposure.

IOS score can be summarised by their mean, SD, median, maximum, IQR, or the 95th percentile value. This documentation describes the method of the simulation, which is based on IOS 1. 

We assigned the calculated IOS scores to random variant and tested whether the hetorogeneity of the IVW MR model still decreases. 100 permutations were performed for each summary measure of IOS 1 (e.g. mean). The script returns the result of IVW Radial MR analysis including the estimates, Q statistics, and p values. 


```{r}
#------------------------------------------------------------------------
#Method: modified second order * IOS_mean
#------------------------------------------------------------------------
ios_dat <-ios(exp=exp_dat, bg=bg_dat)

a <- list()
invisible(capture.output(a <- lapply(1:100, function(x) {

  #IOS scores are assigned to randome SNP
  ios_dat$SNP <- sample(ios_dat$SNP)
  ios_radialMR <- mr.ios_type(dat=dat, ios = ios_dat, ios_type="ios1_sd", alpha = 0.05, 6, tol = 0.0001)

  temp <- ios_radialMR$coef[1, ]
  temp <- tibble::rownames_to_column(temp, "method")
  temp$exposure <- dat$exposure[1]
  temp$outcome <- dat$outcome[1]
  temp$nsnp <- length(dat$SNP)
  temp$Q <- ios_radialMR$qstatistic
  temp$Q_pval <- Total_Q_chi<-pchisq(ios_radialMR$qstatistic, ios_radialMR$df, lower.tail = FALSE)

  temp2 <- temp[, c(6, 7, 1, 8, 2, 3, 4, 5, 9, 10)]

  return(temp2)
}
))
)

a <- data.frame(t(sapply(a, c)))
a$permutation <- rownames(a)

temp <- orig.est(dat)
set1 <- bind.est(temp, a)

set1$method <- "Mod.2nd * ios1_mean"


#------------------------------------------------------------------------
#Method: modified second order * IOS1_sd
#------------------------------------------------------------------------

ios_dat <-ios(exp=exp_dat, bg=bg_dat)

b <- list()
invisible(capture.output(b <- lapply(1:100, function(x) {

  ios_dat$SNP <- sample(ios_dat$SNP)
  ios_radialMR <- mr.ios_type(dat=dat, ios = ios_dat, ios_type="ios1_sd", alpha = 0.05, 6, tol = 0.0001)
  temp <- ios_radialMR$coef[1, ]
  temp <- tibble::rownames_to_column(temp, "method")
  temp$exposure <- dat$exposure[1]
  temp$outcome <- dat$outcome[1]
  temp$nsnp <- length(dat$SNP)
  temp$Q <- ios_radialMR$qstatistic
  temp$Q_pval <- Total_Q_chi<-pchisq(ios_radialMR$qstatistic, ios_radialMR$df, lower.tail = FALSE)

  temp2 <- temp[, c(6, 7, 1, 8, 2, 3, 4, 5, 9, 10)]

  return(temp2)
}
))
)

b <- data.frame(t(sapply(b, c)))
b$permutation <- rownames(b)


temp <- orig.est(dat = dat, ios_type = "ios1_sd")
set2 <- bind.est(temp, b)

set2$method <- "Mod.2nd * ios1_sd"


#------------------------------------------------------------------------
#Method: modified second order * IOS1_iqr
#------------------------------------------------------------------------

ios_dat <-ios(exp=exp_dat, bg=bg_dat)

c <- list()
invisible(capture.output(c <- lapply(1:100, function(x) {

  ios_dat$SNP <- sample(ios_dat$SNP)
  ios_radialMR <- mr.ios_type(dat=dat, ios = ios_dat, ios_type="ios1_iqr", alpha = 0.05, 6, tol = 0.0001)
  temp <- ios_radialMR$coef[1, ]
  temp <- tibble::rownames_to_column(temp, "method")
  temp$exposure <- dat$exposure[1]
  temp$outcome <- dat$outcome[1]
  temp$nsnp <- length(dat$SNP)
  temp$Q <- ios_radialMR$qstatistic
  temp$Q_pval <- Total_Q_chi<-pchisq(ios_radialMR$qstatistic, ios_radialMR$df, lower.tail = FALSE)

  temp2 <- temp[, c(6, 7, 1, 8, 2, 3, 4, 5, 9, 10)]

  return(temp2)
}
))
)

c <- data.frame(t(sapply(c, c)))
c$permutation <- rownames(c)


temp <- orig.est(dat = dat, ios_type = "ios1_iqr")
set3 <- bind.est(temp, c)

set3$method <- "Mod.2nd * ios1_iqr"


#------------------------------------------------------------------------
#Method: modified second order * IOS1_median
#------------------------------------------------------------------------
ios_dat <-ios(exp=exp_dat, bg=bg_dat)

d <- list()
invisible(capture.output(d <- lapply(1:100, function(x) {

  ios_dat$SNP <- sample(ios_dat$SNP)
  ios_radialMR <- mr.ios_type(dat=dat, ios = ios_dat, ios_type="ios1_median", alpha = 0.05, 6, tol = 0.0001)
  temp <- ios_radialMR$coef[1, ]
  temp <- tibble::rownames_to_column(temp, "method")
  temp$exposure <- dat$exposure[1]
  temp$outcome <- dat$outcome[1]
  temp$nsnp <- length(dat$SNP)
  temp$Q <- ios_radialMR$qstatistic
  temp$Q_pval <- Total_Q_chi<-pchisq(ios_radialMR$qstatistic, ios_radialMR$df, lower.tail = FALSE)

  temp2 <- temp[, c(6, 7, 1, 8, 2, 3, 4, 5, 9, 10)]

  return(temp2)
}
))
)

d <- data.frame(t(sapply(d, c)))
d$permutation <- rownames(d)

temp <- orig.est(dat = dat, ios_type = "ios1_median")
set4 <- bind.est(temp, d)

set4$method <- "Mod.2nd * ios1_median"

#------------------------------------------------------------------------
#Method: modified second order * IOS1_95
#------------------------------------------------------------------------

ios_dat <-ios(exp=exp_dat, bg=bg_dat)

e <- list()
invisible(capture.output(e <- lapply(1:100, function(x) {

  ios_dat$SNP <- sample(ios_dat$SNP)
  ios_radialMR <- mr.ios_type(dat=dat, ios = ios_dat, ios_type="ios1_95", alpha = 0.05, 6, tol = 0.0001)
  temp <- ios_radialMR$coef[1, ]
  temp <- tibble::rownames_to_column(temp, "method")
  temp$exposure <- dat$exposure[1]
  temp$outcome <- dat$outcome[1]
  temp$nsnp <- length(dat$SNP)
  temp$Q <- ios_radialMR$qstatistic
  temp$Q_pval <- Total_Q_chi<-pchisq(ios_radialMR$qstatistic, ios_radialMR$df, lower.tail = FALSE)

  temp2 <- temp[, c(6, 7, 1, 8, 2, 3, 4, 5, 9, 10)]

  return(temp2)
}
))
)

e <- data.frame(t(sapply(e, c)))
e$permutation <- rownames(e)

temp <- orig.est(dat = dat, ios_type = "ios1_95")
set5 <- bind.est(temp, e)

set5$method <- "Mod.2nd * ios1_95"

#------------------------------------------------------------------------
#Method: modified second order * IOS_max
#------------------------------------------------------------------------

ios_dat <-ios(exp=exp_dat, bg=bg_dat)

f <- list()
invisible(capture.output(f <- lapply(1:100, function(x) {

  ios_dat$SNP <- sample(ios_dat$SNP)
  ios_radialMR <- mr.ios_type(dat=dat, ios = ios_dat, ios_type="ios1_max", alpha = 0.05, 6, tol = 0.0001)
  temp <- ios_radialMR$coef[1, ]
  temp <- tibble::rownames_to_column(temp, "method")
  temp$exposure <- dat$exposure[1]
  temp$outcome <- dat$outcome[1]
  temp$nsnp <- length(dat$SNP)
  temp$Q <- ios_radialMR$qstatistic
  temp$Q_pval <- Total_Q_chi<-pchisq(ios_radialMR$qstatistic, ios_radialMR$df, lower.tail = FALSE)

  temp2 <- temp[, c(6, 7, 1, 8, 2, 3, 4, 5, 9, 10)]

  return(temp2)
}
))
)

f <- data.frame(t(sapply(f, c)))
f$permutation <- rownames(f)

temp <- orig.est(dat = dat, ios_type = "ios1_max")
set6 <- bind.est(temp, f)

set6$method <- "Mod.2nd * ios1_max"


#------------------------------------------------------------------------
#Generate a dataframe of simulation results
#------------------------------------------------------------------------
ios1_sim <- rbind(set1, set2, set3, set4, set5, set6)
ios1_sim <- as.data.frame(lapply(ios1_sim, unlist))


#Define the class if it is not specified (null)
factor_cols <- c("exposure","outcome", "method")
ios1_sim[factor_cols] <- lapply(ios1_sim[factor_cols], as.factor)
numeric_cols <- c("nsnp", "Estimate", "Std.Error", "t.value", "P_val", "Q", "Q_pval", "permutation")
ios1_sim[numeric_cols] <- lapply(ios1_sim[numeric_cols], as.numeric)

```


The plot presents the Q statistics for MR results. The highlighed dot (red circle) indicates the original Q statistics esimated using correct IOS score. The red circle is expected to be in the middle of the line.

```{r}
#------------------------------------------------------------------------
#Plots
#------------------------------------------------------------------------

high <- ios1_sim %>%
  group_by(method) %>%
  arrange(Q) %>%
  mutate(id = row_number()) %>%
  filter(permutation < 1)

p1 <- ios1_sim %>%
  group_by(method) %>%
  arrange(Q) %>%
  mutate(id = row_number()) %>%
  mutate(highperm = permutation == 0) %>%
  #filter(grepl("mean", method)) %>%
  ggplot(aes(y=Q, x=id, group=as.factor(method)), colour=as.factor(method)) +
  geom_point(aes(colour=as.factor(method)), alpha = 0, shape = ".") +
  geom_line(aes(colour=as.factor(method))) +
  geom_point(data=high, aes(y=Q, x=id), colour='red', size=1) +
  labs(x="Permutations", y="Q statistics", colour="Method")
p1
```
