---
title: "MR.IOS examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


#Empirical examples using mr.ios packages
Here we tested the associations of the following phenotypes;
   1. BMI - CHD
   2. SBP - CHD
   3. EDU - BMI
   4. Sleep - Schizophrenia 
   5. Urate - CHD
   6. Urate - eGFR
   7. CRP - CHD
   
   
```{r}
#install.packages("devtools")
#devtools::install_github("universe77/mr.ios")
library(mr.ios)
library(RadialMR)
library(TwoSampleMR)
library(reshape2)
```


##BMI-CHD
```{r}
# choose exposure - BMI
id_exp <- "ieu-a-2"
exp_dat <- extract_instruments(id_exp)

# choose outcome - CHD
id_out <- "ieu-a-7"
out_dat <- extract_outcome_data(exp_dat$SNP, id_out)

# choose background dataset list
id_bg <- background_ids(id_exp=id_exp, id_out=id_out, type="default")

# make background dataset
bg_dat <- make_background(exp = exp_dat, id_bg = id_bg)

# generate ios
ios_dat <- ios(exp=exp_dat, bg=bg_dat)

# perform mr
dat <- harmonise_data(exp_dat, out_dat)
mr_res <- mr.ios(dat=dat, ios = ios_dat, alpha = 0.05, 1, tol = 0.0001) 
```





##SBP-CHD
# choose exposure - SBP
id_exp <- "ukb-b-20175"
exp_dat <- extract_instruments(id_exp)

# choose outcome - CHD
id_out <- "ieu-a-7"
out_dat <- extract_outcome_data(exp_dat$SNP, id_out)

# choose background dataset list
id_bg <- background_ids(id_exp=id_exp, id_out=id_out, type="default")

# make background dataset
bg_dat <- make_background(exp = exp_dat, id_bg = id_bg)

# generate ios
ios_dat <- ios(exp=exp_dat, bg=bg_dat)

# perform mr
dat <- harmonise_data(exp_dat, out_dat)
mr_res <- mr.ios(dat=dat, ios = ios_dat, alpha = 0.05, 1, tol = 0.0001) 


