---
title: "MR.IOS examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Empirical examples using mr.ios packages
Here we tested the associations of the following phenotypes;
   1. BMI - CHD
   2. SBP - CHD
   3. EDU - BMI
   4. Sleep - Schizophrenia 
   5. Urate - CHD
   6. Urate - eGFR
   7. CRP - CHD
   8. CRP - SCZ
   
   
   
```{r}
#install.packages("devtools")
#devtools::install_github("universe77/mr.ios")
#devtools::install_github("universe77/RadialMR")

library(devtools)
library(mr.ios)
library(RadialMR)
library(TwoSampleMR)
library(dplyr)
library(tidyverse)
```


## SBP-CHD

```{r}
# choose exposure - SBP
id_exp <- "ukb-b-20175"
exp_dat <- extract_instruments(id_exp)

# choose outcome - CHD
id_out <- "ieu-a-7"
out_dat <- extract_outcome_data(exp_dat$SNP, id_out)

# choose background dataset list
id_bg <- background_ids(id_exp=id_exp, id_out=id_out, type = "broad")
id_bg2 <- background_ids(id_exp=id_exp, id_out=id_out, type = "default")

# make background dataset
bg_dat <- make_background(exp = exp_dat, id_bg = id_bg)
bg_dat2 <- make_background(exp = exp_dat, id_bg = id_bg2)
```

For IVW-Radial MR analysis using IOS as an external weight:
```{r}
#----------------------------------------------------------------------------------------------
# MR IOS
#----------------------------------------------------------------------------------------------
# generate ios
ios_dat <- ios(exp=exp_dat, bg=bg_dat)
ios_dat2 <- ios(exp=exp_dat, bg=bg_dat2)

# perform mr
dat <- harmonise_data(exp_dat, out_dat)

res <- list()
res[[1]] <- mr.ios(dat=dat, ios = ios_dat2, ios_type="ios1_mean", alpha = 0.05, weights = 6, tol = 0.0001)
res[[2]] <- mr.ios(dat=dat, ios = ios_dat2, ios_type="ios2_mean", alpha = 0.05, weights = 6, tol = 0.0001)

res[[3]] <- mr.ios(dat=dat, ios = ios_dat, ios_type="ios1_mean", alpha = 0.05, weights = 6, tol = 0.0001)
res[[4]] <- mr.ios(dat=dat, ios = ios_dat, ios_type="ios2_mean", alpha = 0.05, weights = 6, tol = 0.0001)

```


For MR in each group of cluster for instruments:
```{r}
#----------------------------------------------------------------------------------------------
# MR Clustering
#----------------------------------------------------------------------------------------------
kmean <- kmeans_instruments(bg_dat, value_column = "rsq.outcome", kmax=15, nstart=50, iter.max=15)
hclust <- hclust_instruments(bg_dat, value_column = "rsq.outcome", kmax = 50)

# Add info of clustering to harmonised data
dat <- merge(dat, kmean, by = "SNP")

res_clust <- list()
for (i in max(dat[ , ncol(dat)]))
   {
      temp <- subset(dat, dat[ , ncol(dat)] == i)  
      dat_rmr <- RadialMR::format_radial(dat$beta.exposure, dat$beta.outcome, dat$se.exposure, dat$se.outcome, dat$SNP)
      rares <- RadialMR::ivw_radial(dat_rmr, alpha = 0.05, weights = 3, tol = 0.0001, external_weight = F)
      res_clust[[i]] <- rares
   }

```

- Test if esimates from each cluster give substantially different results (durban wu houseman test?)
- Identify the background traits related to each cluster
